### 2019-10-22 README.Rmd
added abstract draft

## Objectives

## Design

## Setting

England and Wales.

## Outcome measures

@@

@@

>>>>

## Background

The Enhanced Tuberculosis Surveillance (ETS) system is a routine surveillance system that collects data on all notified tuberculosis (TB) cases in England and is routinely used in research. Routine data often has a large amount of missing data which may introduce bias to analyses. This study explores the evidence for associations between missingness in several outcomes and demographic variables. Any such associations may introduce bias if not accounted for.

## Methods

We used data on all notifications in the ETS for England (2000-2015). We summarised the proportion of missing data and then used logistic regression to explore the associations between missingness and demographic variables for the following outcomes: drug resistance; BCG status and year; date of symptom onset, diagnosis, notification, starting treatment, finishing treatment and death; and cause of death. For all date variables, we visualised the distribution annually and by month, identifying variables at risk of bias.

@@

All demographic variables considered were associated with data being missing for multiple outcomes. Missingness was not associated with all demographic variables for all outcomes. Associations that were present did not all have the same direction of effect. We found that the date of symptom onset and the date of ending treatment had a high proportion of cases occurring in January and on the 1st and 14th of each month indicating the presence of potential bias.

@@

Missingness in outcomes was associated with multiple demographic variables. These associations could not be generalised, preventing a systematic approach to dealing with missingness. The identified associations may induce bias, meaning that complete case analysis should not be used for this data source - or other similar surveillance data. Our findings should be used motivate the use of imputation and to inform the specification of imputation models for similar surveillance data.

## Keywords

Tuberculosis, surveillance, missingness, imputation
